Preparing your manuscript

### 2025-02-05 article/article.qmd
fix typos and grammar

    - Be set as a single paragraph.
    - Be less than 250 words for all journals except GRL, for which the limit is 150 words.
    - Not include table or figure mentions.
    - Avoid reference citations unless dependent on or directly related to another paper (e.g., companion, comment, reply, or commentary on another paper(s)). AGU’s Style Guide discusses formatting citations in abstracts.
    - Define all abbreviations.
plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: bibliography.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
execute: 
  cache: true
  echo: false
  message: false
  warning: false
---

@@

Cold fronts are a important feature of weather in extratropical regions and are in part responsible for the precipitation in these regions [@konstali2024]. Cold fronts are also linked to extreme events like heavy precipitation [@catto2013] and severe wildfires in southern Australia [@cai2022].

Front identification can be a complex task, to this date, there are several methods that uses different characteristics of the fronts. Some of them looks for thermodynamical changes associated with temperature and humidity in different atmospheric levels [@berry2011], others identify changes in the wind field [@simmonds2011]. These methods relies on the data being in a regular grid (as outputs from numerical models or reanalysis) and usually depends on different thresholds that need to be calibrated for different model resolutions. 
For this study we use the changes from day to day in near surface temperature (or drops in temperature) to estimate the passage of fronts [@reeder2015]. This very simple method can be use for both data from station observations and gridded data (with varied resolution up to 2.5 degrees [@king2024]). Previous studies have shown that this method is suitable for the detection of strong cold fronts in South Eastern Australia [@cai2022, @reader2015], and recently @king2024 showed that this method can successfully identify cold fronts with similar results as the obtained with other methodologies. Moreover, @king2024 calculated a climatology of front frequency using global climate models from phase 6 of the Coupled Model Intercomparison Project (CMIP6). while the models overestimate the frequency of fronts in polar regions and tend to underestimate the occurrence of large temperature drops over the midlatitude land regions compared with ERA5. However, they can represent well the patterns of front frequency equatorward of 60$ ^{\circ}$.

As the planet keeps heating up, understanding how the increased global temperature will impact the occurrence of cold fronts is very important. A few authors have aproached this question, for example @catto2014 used objective front identification [@berry2011] in 18 CMIP5 models and compared the changes between historical 1980-2005 + RCP8.5 2080-2100. They found an overall decrease in front frequency in the Northern Hemisphere, with a poleward shift of the maxima of front frequency and a strong decrease at high latitudes. The Southern Hemisphere shows a poleward shift of the frequency maximum. On the other hand, @konstali2024, identified fronts using gradient of equivalent potential temperature in 850 hPa for the CESM2-LE model and found an increase in the frequency of fronts related to fronts becoming larger or more elongated in both hemispheres.

In this paper we build on the drops in temperature methodology to evaluate the change in frequency of cold fronts in a warmer future compared with the resent past. In an effort to understand possible causes to the potential changes, we explored the relationship between changes in front frequency and transient eddy kinetic energy, as a proxy for storm tracks. Several studies have shown a poleward shift on the storm tracks, particularly in the Southern Hemisphere [@ogorman2010, @tamarin2017, @chenke2022], and these changes could potentially influence the location and intensity of future cold fronts.

>>>>

    - Be set as a single paragraph.
    - Be less than 250 words for all journals except GRL, for which the limit is 150 words.
    - Not include table or figure mentions.
    - Avoid reference citations unless dependent on or directly related to another paper (e.g., companion, comment, reply, or commentary on another paper(s)). AGU’s Style Guide discusses formatting citations in abstracts.
    - Define all abbreviations.
plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: cold-fronts.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
execute: 
  cache: true
  echo: false
  message: false
  warning: false
---

@@

Cold fronts are a important feature of weather in extratropical regions and are partly responsible for the precipitation in these areas [@konstali2024]. Cold fronts are also linked to extreme events, such as heavy precipitation [@catto2013] and severe wildfires in southern Australia [@cai2022].

Identifying fronts can be a complex task. To date, several computational methods have been developed that use different characteristics of fronts. Some of these methods looks for thermodynamic changes associated with temperature and humidity in different atmospheric levels [@berry2011], others identify changes in the wind field [@simmonds2012]. These methods rely on the data being in a regular grid (e.g. outputs from numerical models or reanalysis) and it results depend on setting different thresholds that need to be calibrated for different model resolutions.

For this study we use day-to-day changes in near-surface temperature (or drops in temperature) to estimate the passage of fronts [@reeder2015]. This very simple method can be applied to both data from station observations and gridded data (with varying resolution, up to 2.5 degrees [@king2024]). Previous studies have shown that this method is suitable for the detection of strong cold fronts in southeastern Australia [@cai2022, @reeder2015]. Recently, @king2024 demonstrated that this method can successfully identify cold fronts with results comparable to those obtained with other methodologies. Furthermore, @king2024 calculated a climatology of front frequency using global climate models from phase 6 of the Coupled Model Intercomparison Project (CMIP6). While the models tend to overestimate front frequency in polar regions and underestimate the occurrence of large temperature drops over midlatitude land regions compared with ERA5, they accurately represent the patterns of front frequency equatorward of 60°.

As global temperatures continue to rise, understanding how this warming will impact the occurrence of cold fronts is very important. A few studies have addressed this question, for example @catto2014 used objective front identification [@berry2011] in 18 CMIP5 models and compared the changes between historical simulations between 1980-2005 and the Representative Concentration Pathway 8.5 between 2080-2100. They found an overall decrease in front frequency in the Northern Hemisphere, with a poleward shift in the maximum frequency and a significant decrease at high latitudes. In contrast, the Southern Hemisphere showed a poleward shift in the frequency maximum. On the other hand, @konstali2024, identified fronts in the CESM2-LE model simulations using gradient of equivalent potential temperature in 850 hPa and found an increase in the frequency of fronts that could be related to fronts becoming larger or more elongated in both hemispheres.

In this paper we build on the drops in temperature methodology to evaluate the change in frequency of cold fronts in a warmer future compared with the resent past using a set of models from CMIP6. In an effort to understand possible causes to the potential changes, we explored the relationship between changes in front frequency and transient eddy kinetic energy, that is directly related to storm tracks. Several studies have shown a poleward shift on the storm tracks in future climate change scenarios, particularly in the Southern Hemisphere [@ogorman2010, @tamarin2017, @chemke2022], and these changes could potentially influence the location and intensity of future cold fronts.
